Opportunistic attachment assembles plant-pollinator networks: A walk through of the analysis

### 2017-05-18 analysis/writeup/writeup.tex
lots of small cahnges for reproducibility

In our study we examining the temporal dynamics of plant-pollinator
network assembly using variety of different methods include 1) network
change point detection 2) node/species-level position variation, 3)
species and interaction turnover 4) network-level metrics and 5)
extinction simulations. We are committed to reproducible science and
all analytical code will be maintained on github, along with this
write up.

The entire analysis is executable from the main.sh file. All of the
packages needed to run the analyses are listed in the packages.sh
file, including the specific versions of the python dendropy library,
which is very finicky. We have had no problems with back compatibility
for the R scripts (in terms of R versions from v.~2-3.3 or packages)
but the change point analysis must be run in python2 (not easily
convertible to python3, though this is potentially in progress).

@@

We will walk through each analysis individually.

@@

The output of the change point analysis is saved as a list of the
change points identified and like likelihood scores (formatted by
prepChangePointOutput.R).

@@

In this analysis the assembling hedgerows are set as the intercept, to
weedy field margins ("control") have significantly less change points
than assembling hedgerows ("maturing"). Mature and maturing are about
the same.

@@

We can then use a linear mixed model to regress the cv of weighted
closeness by rarefied degree and persistence. Random effects of site
and species are also included.

@@

We see the same pattern in the plant data.

@@

\hlcom{## occ.date is significant}
\hlstd{pol.lm.cv.quant} \hlkwb{<-} \hlkwd{lqmm}\hlstd{(}\hlkwc{fixed}\hlstd{=cv} \hlopt{~} \hlstd{occ.date} \hlopt{+} \hlstd{r.degree,} \hlkwc{random}\hlstd{=}\hlopt{~}\hlnum{1}\hlstd{,}
                        \hlkwc{group}\hlstd{= GenusSpecies,}
                        \hlkwc{data}\hlstd{=pol.cv}\hlopt{$}\hlstd{lm.data,}
                        \hlkwc{control}\hlstd{=}\hlkwd{list}\hlstd{(}\hlkwc{LP_max_iter}\hlstd{=}\hlnum{10}\hlopt{^}\hlnum{4}\hlstd{))}
\hlstd{pol.sum.boot.quant} \hlkwb{<-} \hlkwd{summary.boot.lqmm}\hlstd{(}\hlkwd{boot}\hlstd{(pol.lm.cv.quant,}
                                             \hlkwc{R}\hlstd{=}\hlnum{100}\hlstd{))}
\end{alltt}
\begin{verbatim}
## Quantile 0.5 
##                  Value Std. Error lower bound upper bound  Pr(>|t|)    
## (Intercept)  3.7160065  0.0920970   3.5332661      3.8987 < 2.2e-16 ***
## occ.date     0.7040012  0.2676157   0.1729936      1.2350  0.009884 ** 
## r.degree    -0.0010938  0.0032020  -0.0074472      0.0053  0.733375    
## scale        0.2869339  0.0191323   0.2489713      0.3249 < 2.2e-16 ***
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

Next, in the function calcBetaStatus.R executed by beta-div.R, for the
observed and nulls communities, we calculated the pair-wise
dissimilarity between sites.  We then used the null community
dissimilarities to calculate the expected $\beta$-diversity when
communities are randomly assembled but constrained so that they have
either the same 1) number of individuals or 2) species richness as the
observed communities and with species drawn from a meta-community with
the same species abundance distributions.  In order to do this, we
followed \cite{chase-2011-art24}.  Specifically, we calculated the
fraction of randomly assembled communities with dissimilarity values
less than (and half of those equal to) that of the observed community.
We used this fraction as a ``corrected dissimilarity score'' for our
observed data.  Corrected dissimilarity values near one indicate that
our observed communities exhibit more species turnover between sites
than expected under a random assembly process while values near $0.5$
indicate that our observed communities exhibit levels of turnover more
in line with the null expectation. We calculated the corrected
dissimilarities for each type of randomized community. We then
generated principle coordinate axes (PCoA) based on the corrected
pair-wise dissimilarities \citep{oksanen-2013, anderson-2006-683,
anderson-2011-19, anderson-2006-245}.  We calculated dispersion for
each site type by finding the centroid in PCoA space for that site
type and then calculating the distances from sites of that type to
that centroid. The centroid is the point that minimizes the sum of
these distances.

@@

Toggling alpha and beta in the beta-div.R script changes between the
null model options. When binary is TRUE, the binary null models are
used and the dissimilarity method is "jaccard" (all set up in the
initialize\_beta.R). When alpha is TRUE, species richness but not
individual abundances are constrained, and the reverse when alpha is
FALSE. The chao dissimilarity measure is used in both
abundance-weighted cases.

@@

@@

There is little difference between any of the options, except mature
sites have significantly less turnover in pollinators with the alpha
null options, which is likely most appropriate.

Changing "type" to "pols", "plants", or "Ints" calculates turnover of
pollinators, plants and interactions respectively. All of these are
executed in the main.sh script.

We also included a linear model that took temporal auto-correlation
into account, with the same results but not a large gain in fit given
the additional parameters (tested using AIC)

@@

To calculate weighted interaction turnover we borrowed concepts from
calculating phylogenetic $\beta$-diversity. We began by creating a
"phylogenetic" tree of interactions based on their position in the
network using the getLinkCommunities function in the linkcomm package
\citep{kalinka2011linkcomm}. For bi-partite networks, the set of
neighbors is used to count nodes for the edge similarity metric
because node i and node j cannot share an edge in a bi-partite
network. The partition density for bi-partite networks is calculated
as:

@@

where M is the total number of edges, $m_c$ is the number of edges in
subset $c$, $n_c$ is the number of nodes in subset c, $n_{c_0}$ is the
number of nodes in partition 0, and $n_{c_1}$ is the number of nodes
in partition 1.

Use the resulting dendrogram and interaction dissimilarity matrices,
we use the comdist function in the picante package
\citep{picante-2010-1463} to calculate expected "phylogenetic"
distance separating two interactions drawn randomly from different
communities. This gives us a measure of interaction turnover weighted
by how interaction similarity. This is all wrapped up in the function
CalcCommDis

@@

We use a linear mixed model to regress weighted interaction turnover
and site type and find no statistically significant differences.

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{phylo.int.mod} \hlkwb{<-} \hlkwd{lmer}\hlstd{(PhyloInt} \hlopt{~} \hlstd{SiteStatus} \hlopt{+}
                          \hlstd{(}\hlnum{1}\hlopt{|}\hlstd{Site),}
                      \hlkwc{data}\hlstd{=phylo.int}\hlopt{$}\hlstd{phylo.int)}
\hlkwd{print}\hlstd{(}\hlkwd{summary}\hlstd{(phylo.int.mod))}
\end{alltt}
\begin{verbatim}
## Linear mixed model fit by REML t-tests use Satterthwaite approximations
##   to degrees of freedom [lmerMod]
## Formula: PhyloInt ~ SiteStatus + (1 | Site)
##    Data: phylo.int$phylo.int
## 
## REML criterion at convergence: -1087.4
## 
## Scaled residuals: 
##     Min      1Q  Median      3Q     Max 
## -4.4271 -0.3718  0.1392  0.5481  2.9588 
## 
## Random effects:
##  Groups   Name        Variance Std.Dev.
##  Site     (Intercept) 0.003341 0.05781 
##  Residual             0.004470 0.06686 
## Number of obs: 459, groups:  Site, 38
## 
## Fixed effects:
##                    Estimate Std. Error       df t value Pr(>|t|)    
## (Intercept)         0.90778    0.01965 37.00000  46.198  < 2e-16 ***
## SiteStatuscontrol  -0.08206    0.02414 35.13000  -3.400  0.00169 ** 
## SiteStatusmaturing  0.03339    0.03065 33.20000   1.089  0.28385    
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
## 
## Correlation of Fixed Effects:
##             (Intr) StSttsc
## StSttscntrl -0.814        
## StSttsmtrng -0.641  0.522
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

The baci.R script (before-after-control-impact, baci, the orginal
design of the study, vs.~ the word for kisses in Italian) calculates
various network metrics for the assembling hedgerows, standardizes
them using null communities, then include then as response variables
in linear models. The number of null communities can be changed by
setting N at the top of the script.

Calculate and standardize the network metrics. CalcNetworkMetrics
takes an argument "networkLevel.mets" which is passed to networkLevel
within the package bipartite \citep{bipartite} if other metrics are
desired. Unlike the turnover analysis, the networks include
non-integers because the mean was taken over sampling rounds within a
year. We therefore use the null model proposed by \citep{Galeano2009}
to randomize the communities while fixing the total number of
interactions, species and interaction frequency distributions (though
any function that takes a single interaction matrix as an argument can
be substituted using the null.fun argument of the calcNullStat
function.

Calculate and pretty network metrics.

@@

Then regress the nestedness, modularity, specialization and
connectance by time ("ypr" in the data, years post restoration).

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlkwd{lapply}\hlstd{(mods, summary)}
\end{alltt}
\end{kframe}
\end{knitrout}

As per the suggestion of reviewers, we also tested the same hypotheses
with linear mixed models that included temporal auto-correlation, and
splines. Like the turnover models, the temporal auto-correlation
models showed similar results, but the fit was not substantially
improved by adding the additional parameters. The spline chose to fit
a line, indicating that the linear models were sufficient. They are
time consuming to run so the results are not included here.

@@

This area is then regressed against time ("ypr" again).

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlcom{## no effect of ypr on robustness}
\hlstd{mod.ypr} \hlkwb{<-} \hlkwd{lmer}\hlstd{(Robustness} \hlopt{~} \hlstd{ypr}
                \hlopt{+} \hlstd{(}\hlnum{1}\hlopt{|}\hlstd{Site)} \hlopt{+} \hlstd{(}\hlnum{1}\hlopt{|}\hlstd{Year),}
                \hlkwc{data}\hlstd{=res[}\hlopt{!}\hlkwd{is.na}\hlstd{(res}\hlopt{$}\hlstd{ypr),])}
\hlkwd{summary}\hlstd{(mod.ypr)}
\end{alltt}
\begin{verbatim}
## Linear mixed model fit by REML t-tests use Satterthwaite approximations
##   to degrees of freedom [lmerMod]
## Formula: Robustness ~ ypr + (1 | Site) + (1 | Year)
##    Data: res[!is.na(res$ypr), ]
## 
## REML criterion at convergence: -167.6
## 
## Scaled residuals: 
##      Min       1Q   Median       3Q      Max 
## -2.03661 -0.73640 -0.02933  0.62103  2.40910 
## 
## Random effects:
##  Groups   Name        Variance  Std.Dev.
##  Site     (Intercept) 0.0019852 0.04456 
##  Year     (Intercept) 0.0001163 0.01078 
##  Residual             0.0048546 0.06967 
## Number of obs: 80, groups:  Site, 14; Year, 8
## 
## Fixed effects:
##              Estimate Std. Error        df t value Pr(>|t|)    
## (Intercept)  0.331489   0.016356 11.040000  20.267 4.36e-10 ***
## ypr          0.001166   0.004144 34.880000   0.281     0.78    
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
## 
## Correlation of Fixed Effects:
##     (Intr)
## ypr -0.395
\end{verbatim}
\end{kframe}
\end{knitrout}

>>>>

In our study we examin the temporal dynamics of plant-pollinator
network assembly using variety of different methods including 1) network
change point detection 2) node/species-level position variation, 3)
species and interaction turnover 4) network-level metrics and 5)
extinction simulations. We are committed to reproducible science and
all analytical code will be maintained on github, along with this
write up.

The entire analysis is executable from the main.sh file. All of the
packages needed to run the analyses are listed in the packages.sh
file, including the specific versions of the python dendropy library,
which is very finicky. We have had no problems with backward
compatibility for the R scripts (in terms of R versions from v.~2-3.4
or packages) but the change point analysis must be run in python2 (not
easily convertible to python3, though this is potentially in
progress).

@@

We will walk through each the main script for each analysis
individually.

@@

The output of the change point analysis is saved as a list of the
change points identified and likelihood scores (formatted by
prepChangePointOutput.R).

@@

In this analysis the assembling hedgerows are set as the intercept, so
weedy field margins ("control") have significantly less change points
than assembling hedgerows ("maturing"). Mature and maturing are about
the same.

@@

We can then use a linear mixed model to regress the cv of weighted
closeness by rarefied degree and persistence. Random effects of site
and species are included. Because degrees of freedom are not
well-dinfed for a mixed model Satterthwaite approximations implemented
in the lmeTest package are used so that p-values can be calculated for
the linear mixed models \citep{lmetest}.

@@

We found no statistically significant relationships in the plant data.

@@

\hlcom{## occ.date is significant}
\hlstd{pol.lm.cv.quant} \hlkwb{<-} \hlkwd{lqmm}\hlstd{(}\hlkwc{fixed}\hlstd{=cv} \hlopt{~} \hlstd{occ.date} \hlopt{+} \hlstd{r.degree,} \hlkwc{random}\hlstd{=}\hlopt{~}\hlnum{1}\hlstd{,}
                        \hlkwc{group}\hlstd{= GenusSpecies,}
                        \hlkwc{data}\hlstd{=pol.cv}\hlopt{$}\hlstd{lm.data,}
                        \hlkwc{control}\hlstd{=}\hlkwd{list}\hlstd{(}\hlkwc{LP_max_iter}\hlstd{=}\hlnum{10}\hlopt{^}\hlnum{4}\hlstd{))}
\hlstd{pol.sum.boot.quant} \hlkwb{<-} \hlkwd{summary.boot.lqmm}\hlstd{(}\hlkwd{boot}\hlstd{(pol.lm.cv.quant,}
                                             \hlkwc{R}\hlstd{=}\hlnum{100}\hlstd{))}
\end{alltt}
\begin{verbatim}
## Quantile 0.5 
##                  Value Std. Error lower bound upper bound Pr(>|t|)    
## (Intercept)  3.7160065  0.0854433   3.5464685      3.8855  < 2e-16 ***
## occ.date     0.7040012  0.2454156   0.2170433      1.1910  0.00504 ** 
## r.degree    -0.0010938  0.0032266  -0.0074960      0.0053  0.73533    
## scale        0.2869339  0.0185467   0.2501333      0.3237  < 2e-16 ***
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

Next, in the function calcBetaStatus.R executed by beta-div.R, for the
observed and nulls communities, we calculated the pair-wise
dissimilarity between sites.  We then used the null community
dissimilarities to calculate the expected $\beta$-diversity when
communities are randomly assembled but constrained so that they have
either the same 1) number of individuals or 2) species richness as the
observed communities and with species drawn from a meta-community with
the same species abundance distributions.  In order to do this, we
followed \cite{chase-2011-art24}.  Specifically, we calculated the
fraction of randomly assembled communities with dissimilarity values
less than (and half of those equal to) that of the observed community.
We used this fraction as a ``corrected dissimilarity score'' for our
observed data.  Corrected dissimilarity values near one indicate that
our observed communities exhibit more species turnover between sites
than expected under a random assembly process while values near $0.5$
indicate that our observed communities exhibit levels of turnover more
in line with the null expectation. We calculated the corrected
dissimilarities for each type of randomized community. We then
generated principle coordinate axes (PCoA) based on the corrected
pair-wise dissimilarities \citep{oksanen-2013, anderson-2006-683,
anderson-2011-19, anderson-2006-245}.  We calculated dispersion for
each site type by finding the centroid in PCoA space for that site
type and then calculating the distances from sites of that type to
that centroid using the betadisper function in the vegan packages
\citep{oksanen-2013}. The centroid is the point that minimizes the sum
of these distances.

@@

Toggling alpha and beta in the beta-div.R script changes between the
null model options. When binary is TRUE, the binary null models are
used and the dissimilarity method is "jaccard" (all set up in the
initialize\_beta.R). When alpha is TRUE, species richness but not
individual abundances are constrained, and the reverse when alpha is
FALSE. The chao dissimilarity measure is used in both
abundance-weighted cases. There is little difference between any of
the options, except mature sites have significantly less turnover in
pollinators with the alpha null options, which is likely most
appropriate.

@@

Changing "type" to "pols", "plants", or "Ints" calculates turnover of
pollinators, plants and interactions respectively. All of these
options are executed in the main.sh script.

@@

Follow a reviwer's suggestion, we also included a linear model that
took temporal auto-correlation into account, with the same results but
not a large gain in fit given the additional parameters (tested using
AIC)

@@

To calculate weighted interaction turnover we borrowed concepts from
calculating phylogenetic $\beta$-diversity. We began by creating a
"phylogenetic" tree of interactions based on their position in the
network using the getLinkCommunities function in the linkcomm package
\citep{kalinka2011linkcomm}. For bipartite networks, the set of
neighbors is used to count nodes for the edge similarity metric
because node i and node j cannot share an edge in a bipartite
network. The partition density for bipartite networks is calculated
as:

@@

where M is the total number of edges, $m_c$ is the number of edges in
subset $c$, $n_c$ is the number of nodes in subset c, $n_{c_0}$ is the
number of nodes in partition 0, and $n_{c_1}$ is the number of nodes
in partition 1 (from the getLinkCommunities helpfile).

With the resulting dendrogram and interaction dissimilarity matrices,
we use the comdist function in the picante package
\citep{picante-2010-1463} to calculate expected "phylogenetic"
distance separating two interactions drawn randomly from different
communities. This gives us a measure of interaction turnover weighted
by how interaction similarity. This is all wrapped up in the function
CalcCommDis

@@

abund.w = TRUE weighted dissimilarity scores by species abundances.

We use a linear mixed model to regress weighted interaction turnover. When
interactions were weighted by their similarity, both assembling and
mature hedgerows had higher rates of turnover than field margins.

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{phylo.int.mod} \hlkwb{<-} \hlkwd{lmer}\hlstd{(PhyloInt} \hlopt{~} \hlstd{SiteStatus} \hlopt{+}
                          \hlstd{(}\hlnum{1}\hlopt{|}\hlstd{Site),}
                      \hlkwc{data}\hlstd{=phylo.int}\hlopt{$}\hlstd{phylo.int)}
\hlkwd{print}\hlstd{(}\hlkwd{summary}\hlstd{(phylo.int.mod))}
\end{alltt}
\begin{verbatim}
## Linear mixed model fit by REML t-tests use Satterthwaite approximations
##   to degrees of freedom [lmerMod]
## Formula: PhyloInt ~ SiteStatus + (1 | Site)
##    Data: phylo.int$phylo.int
## 
## REML criterion at convergence: -1087.4
## 
## Scaled residuals: 
##     Min      1Q  Median      3Q     Max 
## -4.4271 -0.3718  0.1392  0.5481  2.9588 
## 
## Random effects:
##  Groups   Name        Variance Std.Dev.
##  Site     (Intercept) 0.003341 0.05781 
##  Residual             0.004470 0.06686 
## Number of obs: 459, groups:  Site, 38
## 
## Fixed effects:
##                   Estimate Std. Error       df t value Pr(>|t|)    
## (Intercept)        0.94117    0.02352 30.88000  40.014  < 2e-16 ***
## SiteStatuscontrol -0.11545    0.02738 31.12000  -4.216 0.000198 ***
## SiteStatusmature  -0.03339    0.03065 33.20000  -1.089 0.283853    
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
## 
## Correlation of Fixed Effects:
##             (Intr) StSttsc
## StSttscntrl -0.859        
## SiteSttsmtr -0.767  0.659
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

The baci.R script (before-after-control-impact, baci, the orginal
design of the study, vs.~the word for kisses in Italian) calculates
various network metrics for the assembling hedgerows, standardizes
them using null communities, then include then as response variables
in linear models. The number of null communities can be changed by
setting N at the top of the script.

CalcNetworkMetrics takes an argument "index" which is passed to
networkLevel within the package bipartite \citep{bipartite} if other
metrics are desired. Unlike the turnover analysis, the networks
include non-integers because the mean was taken over sampling rounds
within a year. We therefore use the null model proposed by
\citep{Galeano2009} to randomize the communities while fixing the
total number of interactions, species and interaction frequency
distributions (though any function that takes a single interaction
matrix as an argument can be substituted using the null.fun argument
of the calcNullStat function.

@@

Then regress the nestedness, modularity, specialization and
connectance, plant/pollinator species richness by time ("ypr" in the
data, years post restoration).

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlcom{## ************************************************************}
\hlcom{## linear mixed models}
\hlstd{ys} \hlkwb{<-} \hlkwd{c}\hlstd{(}\hlstr{"zNODF"}\hlstd{,} \hlstr{"zmod.met.D"}\hlstd{,} \hlstr{"zH2"}\hlstd{,} \hlstr{"connectance"}\hlstd{)}

\hlstd{formulas} \hlkwb{<-}\hlkwd{lapply}\hlstd{(ys,} \hlkwa{function}\hlstd{(}\hlkwc{x}\hlstd{) \{}
    \hlkwd{as.formula}\hlstd{(}\hlkwd{paste}\hlstd{(x,} \hlstr{"~"}\hlstd{,}
                     \hlkwd{paste}\hlstd{(}\hlstr{"scale(ypr)"}\hlstd{,}
                           \hlstr{"(1|Site)"}\hlstd{,}
                           \hlstr{"(1|Year)"}\hlstd{,}
                           \hlkwc{sep}\hlstd{=}\hlstr{"+"}\hlstd{)))}
\hlstd{\})}

\hlstd{formulas.spac} \hlkwb{<-}\hlkwd{lapply}\hlstd{(ys,} \hlkwa{function}\hlstd{(}\hlkwc{x}\hlstd{) \{}
    \hlkwd{as.formula}\hlstd{(}\hlkwd{paste}\hlstd{(x,} \hlstr{"~"}\hlstd{,}
                     \hlkwd{paste}\hlstd{(}\hlstr{"scale(ypr)"}\hlstd{,}
                           \hlkwc{sep}\hlstd{=}\hlstr{"+"}\hlstd{)))}
\hlstd{\})}

\hlstd{mods} \hlkwb{<-} \hlkwd{lapply}\hlstd{(formulas,} \hlkwa{function}\hlstd{(}\hlkwc{x}\hlstd{)\{}
    \hlkwd{lmer}\hlstd{(x,}
         \hlkwc{data}\hlstd{=baci)}
\hlstd{\})}

\hlkwd{names}\hlstd{(mods)} \hlkwb{<-} \hlstd{ys}
\hlcom{## results}
\hlkwd{lapply}\hlstd{(mods, summary)}
\end{alltt}
\begin{verbatim}
## $zNODF
## Linear mixed model fit by REML t-tests use Satterthwaite approximations
##   to degrees of freedom [lmerMod]
## Formula: zNODF ~ scale(ypr) + (1 | Site) + (1 | Year)
##    Data: baci
## 
## REML criterion at convergence: 186.2
## 
## Scaled residuals: 
##     Min      1Q  Median      3Q     Max 
## -1.7060 -0.5585 -0.1869  0.4507  2.2506 
## 
## Random effects:
##  Groups   Name        Variance Std.Dev.
##  Year     (Intercept) 1.477    1.215   
##  Site     (Intercept) 6.427    2.535   
##  Residual             8.241    2.871   
## Number of obs: 36, groups:  Year, 8; Site, 5
## 
## Fixed effects:
##             Estimate Std. Error     df t value Pr(>|t|)   
## (Intercept)   6.1644     1.3091 4.7280   4.709   0.0061 **
## scale(ypr)    1.2424     0.6717 6.2160   1.850   0.1121   
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
## 
## Correlation of Fixed Effects:
##            (Intr)
## scale(ypr) 0.008 
## 
## $zmod.met.D
## Linear mixed model fit by REML t-tests use Satterthwaite approximations
##   to degrees of freedom [lmerMod]
## Formula: zmod.met.D ~ scale(ypr) + (1 | Site) + (1 | Year)
##    Data: baci
## 
## REML criterion at convergence: 116.8
## 
## Scaled residuals: 
##     Min      1Q  Median      3Q     Max 
## -1.8149 -0.7167 -0.2496  0.7042  2.8513 
## 
## Random effects:
##  Groups   Name        Variance Std.Dev.
##  Year     (Intercept) 0.000    0.000   
##  Site     (Intercept) 0.000    0.000   
##  Residual             1.471    1.213   
## Number of obs: 36, groups:  Year, 8; Site, 5
## 
## Fixed effects:
##             Estimate Std. Error      df t value Pr(>|t|)  
## (Intercept)  -0.4738     0.2022 34.0000  -2.344   0.0251 *
## scale(ypr)    0.2363     0.2050 34.0000   1.152   0.2572  
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
## 
## Correlation of Fixed Effects:
##            (Intr)
## scale(ypr) 0.000 
## 
## $zH2
## Linear mixed model fit by REML t-tests use Satterthwaite approximations
##   to degrees of freedom [lmerMod]
## Formula: zH2 ~ scale(ypr) + (1 | Site) + (1 | Year)
##    Data: baci
## 
## REML criterion at convergence: 189.7
## 
## Scaled residuals: 
##     Min      1Q  Median      3Q     Max 
## -2.0587 -0.4511 -0.1303  0.5563  2.0243 
## 
## Random effects:
##  Groups   Name        Variance Std.Dev.
##  Year     (Intercept) 3.509    1.873   
##  Site     (Intercept) 3.461    1.860   
##  Residual             9.107    3.018   
## Number of obs: 36, groups:  Year, 8; Site, 5
## 
## Fixed effects:
##             Estimate Std. Error      df t value Pr(>|t|)   
## (Intercept)  -5.9515     1.1889  6.2850  -5.006  0.00213 **
## scale(ypr)   -1.4434     0.8571  6.8770  -1.684  0.13680   
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
## 
## Correlation of Fixed Effects:
##            (Intr)
## scale(ypr) 0.025 
## 
## $connectance
## Linear mixed model fit by REML t-tests use Satterthwaite approximations
##   to degrees of freedom [lmerMod]
## Formula: connectance ~ scale(ypr) + (1 | Site) + (1 | Year)
##    Data: baci
## 
## REML criterion at convergence: -99.5
## 
## Scaled residuals: 
##      Min       1Q   Median       3Q      Max 
## -1.62761 -0.54457 -0.04675  0.38328  2.08088 
## 
## Random effects:
##  Groups   Name        Variance  Std.Dev.
##  Year     (Intercept) 0.0000000 0.00000 
##  Site     (Intercept) 0.0009775 0.03126 
##  Residual             0.0021487 0.04635 
## Number of obs: 36, groups:  Year, 8; Site, 5
## 
## Fixed effects:
##              Estimate Std. Error        df t value Pr(>|t|)    
## (Intercept)  0.210108   0.015979  3.937000   13.15 0.000213 ***
## scale(ypr)  -0.022972   0.008004 30.733000   -2.87 0.007359 ** 
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
## 
## Correlation of Fixed Effects:
##            (Intr)
## scale(ypr) -0.003
\end{verbatim}
\end{kframe}
\end{knitrout}

As per the suggestion of reviewers, we also tested the same hypotheses
with linear mixed models that included temporal auto-correlation, and
splines. Like the turnover models, the temporal auto-correlation
models showed similar results, but the fit was not substantially
improved by adding the additional parameters. The spline chose to fit
a line, indicating that the linear models were sufficient.

And the splines:
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlcom{## ************************************************************}
\hlcom{## splines}

\hlcom{## for gaussian models}
\hlstd{mods.splines} \hlkwb{<-} \hlkwd{lapply}\hlstd{(ys, calcSpine)}
\end{alltt}
\begin{verbatim}
## $info
## [1] 0
## 
## $smoothingParameters
##    mu     v 
## 10000 10000 
## 
## $info
## [1] 0
## 
## $smoothingParameters
##    mu     v 
## 11000 10000 
## 
## $info
## [1] 0
## 
## $smoothingParameters
##    mu     v 
## 10000 10000 
## 
## $info
## [1] 0
## 
## $smoothingParameters
##    mu     v 
## 21437  1750
\end{verbatim}
\begin{alltt}
\hlkwd{names}\hlstd{(mods.splines)} \hlkwb{<-} \hlstd{ys}

\hlcom{## for poisson models}
\hlstd{mods.splines.log} \hlkwb{<-} \hlkwd{lapply}\hlstd{(poi.ys, calcSpine,} \hlkwc{calc.log}\hlstd{=}\hlnum{TRUE}\hlstd{)}
\end{alltt}
\begin{verbatim}
## $info
## [1] 0
## 
## $smoothingParameters
##    mu     v 
## 11000 10000 
## 
## $info
## [1] 0
## 
## $smoothingParameters
##    mu     v 
## 11000 10000 
## 
## $info
## [1] 0
## 
## $smoothingParameters
##    mu     v 
## 11000 10000
\end{verbatim}
\begin{alltt}
\hlkwd{names}\hlstd{(mods.splines.log)} \hlkwb{<-} \hlstd{poi.ys}
\end{alltt}
\end{kframe}
\end{knitrout}

The smoothing parameters indicate lines were fit.

@@

This area is then regressed against time ("ypr") Random effects of
year and site are included.

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlcom{## no effect of ypr on robustness}
\hlstd{mod.ypr} \hlkwb{<-} \hlkwd{lmer}\hlstd{(Robustness} \hlopt{~} \hlstd{ypr}
                \hlopt{+} \hlstd{(}\hlnum{1}\hlopt{|}\hlstd{Site)} \hlopt{+} \hlstd{(}\hlnum{1}\hlopt{|}\hlstd{Year),}
                \hlkwc{data}\hlstd{=res[}\hlopt{!}\hlkwd{is.na}\hlstd{(res}\hlopt{$}\hlstd{ypr),])}
\hlkwd{summary}\hlstd{(mod.ypr)}
\end{alltt}
\begin{verbatim}
## Linear mixed model fit by REML t-tests use Satterthwaite approximations
##   to degrees of freedom [lmerMod]
## Formula: Robustness ~ ypr + (1 | Site) + (1 | Year)
##    Data: res[!is.na(res$ypr), ]
## 
## REML criterion at convergence: -166.9
## 
## Scaled residuals: 
##      Min       1Q   Median       3Q      Max 
## -1.99009 -0.62116 -0.07289  0.56358  2.37929 
## 
## Random effects:
##  Groups   Name        Variance  Std.Dev.
##  Site     (Intercept) 2.395e-03 0.048943
##  Year     (Intercept) 5.091e-05 0.007135
##  Residual             4.842e-03 0.069583
## Number of obs: 80, groups:  Site, 14; Year, 8
## 
## Fixed effects:
##              Estimate Std. Error        df t value Pr(>|t|)    
## (Intercept)  0.333299   0.016943 10.950000  19.671 6.85e-10 ***
## ypr          0.000965   0.004133 32.430000   0.233    0.817    
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
## 
## Correlation of Fixed Effects:
##     (Intr)
## ypr -0.374
\end{verbatim}
\end{kframe}
\end{knitrout}

The next estimate of resilience we use is based on how quickly
information spreads through a network as estimated by algebraic
connectivity.
